### 2024-04-04 thesis/sections/6.1-object-identification.tex
learning curves

The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
A random guess corresponds to 50\% in the \emph{Dale-2} dataset, 20\% in the \emph{Dale-5} dataset and 13\% in the \emph{CLEVR color} dataset.

@@

\begin{figure}[ht!]
    \centering
    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-2_vocab-size.png}
        \label{fig:learning-curve_discriminator_dale-2_vocab-size}
    }
    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-2_max-len.png}
        \label{fig:learning-curve_discriminator_dale-2_max-len}
    }
    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-5_vocab-size.png}
        \label{fig:learning-curve_discriminator_dale-5_vocab-size}
    }
    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-5_max-len.png}
        \label{fig:learning-curve_discriminator_dale-5_max-len}
    }
    \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_colour_vocab-size.png}
        \label{fig:learning-curve_discriminator_colour_vocab-size}
    }
    \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_colour_max-len.png}
        \label{fig:learning-curve_discriminator_colour_max-len}
    }
    \caption{Learning curves of all discrimination games on each dataset. In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$. The baseline is marked in black.}
    \label{fig:learning-curves_discriminator}
\end{figure}

When looking at the results of the experiments, it can be seen that the agents are successfully discriminating the objects across all datasets.
Tables \ref{tab:results_discriminator_dale-2} to \ref{tab:results_discriminator_color} give an overview over the accuracy scores of selected configurations.
For each dataset there are many configurations that outperform the baseline models by a high margin.
However, there are still clear differences between the accuracy scores of each dataset.

\cmtDK[inline]{learning curve}
% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge. As David pointed out it may be that convergence is unstable and changes happen in between.
% DK: TODO

\begin{table}[ht]
    \centering
    \begin{tabular}{cccc|c}
        \toprule
                                      &           &     &       & \multicolumn{1}{c}{\textbf{Dale-2}} \\  \cmidrule(lr){5-5}
        $h_r$                         & $h_s$     & $n$ & $|V|$ & \textbf{Accuracy}                   \\\midrule
        {100}                         & {100}     & {2} & {10}  & {99,96\%}                           \\
        {100}                         & {1000}    & {2} & {10}  & {99,96\%}                           \\
        {1000}                        & {100}     & {3} & {10}  & {99,87\%}                           \\
        {500}                         & {100}     & {3} & {100} & {99,83\%}                           \\
        {100}                         & {100}     & {6} & {50}  & {99,83\%}                           \\
        {100}                         & {500}     & {3} & {16}  & {99,7\%}                            \\
        {1000}                        & {100}     & {6} & {16}  & {99,61\%}                           \\
        {100}                         & {100}     & {6} & {100} & {99,57\%}                           \\
        {100}                         & {500}     & {1} & {100} & {99,52\%}                           \\
        {100}                         & {100}     & {1} & {50}  & {99,35\%}                           \\
        {500}                         & {100}     & {1} & {2}   & {99,26\%}                           \\
        {100}                         & {500}     & {1} & {50}  & {99,26\%}                           \\
        {500}                         & {500}     & {1} & {100} & {99,26\%}                           \\
        {100}                         & {500}     & {3} & {100} & {99,83\%}                           \\
        {500}                         & {1000}    & {1} & {16}  & {98\%}                              \\
        {100}                         & {500}     & {1} & {10}  & {96,61\%}                           \\
        {100}                         & {1000}    & {1} & {100} & {99,44\%}                           \\
        {500}                         & {500}     & {6} & {2}   & {95,92\%}                           \\
        {100}                         & {1000}    & {4} & {2}   & {94,88\%}                           \\
        {100}                         & {1000}    & {6} & {10}  & \textcolor{purple}{63,15\%}         \\
        {1000}                        & {500}     & {2} & {2}   & \textcolor{purple}{62,98\%}         \\\midrule
        \multicolumn{4}{c|}{baseline} & {62,72\%}                                                     \\\midrule
        {100}                         & {1000}    & {3} & {2}   & \textcolor{purple}{61,2\%}          \\
        \bottomrule
    \end{tabular}
    \caption{Selected results of the object identification games on the Dale-2 dataset with $e_s=100$ and $e_r=500$}
    \label{tab:results_discriminator_dale-2}
\end{table}

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{cccc|c}
        \toprule
                                      &           &     &       & \multicolumn{1}{c}{\textbf{Dale-5}} \\  \cmidrule(lr){5-5}
        $h_r$                         & $h_s$     & $n$ & $|V|$ & \textbf{Accuracy}                   \\\midrule
        {1000}                        & {500}     & {3} & {50}  & {91,75\%}                           \\
        {100}                         & {100}     & {6} & {10}  & {91,02\%}                           \\
        {100}                         & {100}     & {3} & {10}  & {90,76\%}                           \\
        {100}                         & {100}     & {4} & {16}  & {89,76\%}                           \\
        {100}                         & {100}     & {2} & {10}  & {88,85\%}                           \\
        {500}                         & {1000}    & {1} & {100} & {87,11\%}                           \\
        {1000}                        & {1000}    & {6} & {50}  & {85,07\%}                           \\
        {500}                         & {1000}    & {1} & {10}  & {84,46\%}                           \\
        {500}                         & {500}     & {1} & {50}  & {83,29\%}                           \\
        {100}                         & {100}     & {1} & {10}  & {80,86\%}                           \\
        {500}                         & {100}     & {1} & {100} & {76,74\%}                           \\
        {500}                         & {500}     & {4} & {2}   & {77,78\%}                           \\
        {1000}                        & {1000}    & {3} & {2}   & {75,78\%}                           \\
        {100}                         & {100}     & {3} & {2}   & {74\%}                              \\
        {500}                         & {500}     & {1} & {2}   & {63,41\%}                           \\\midrule
        \multicolumn{4}{c|}{baseline} & {35,07\%}                                                     \\\midrule
        {500}                         & {1000}    & {2} & {100} & \textcolor{purple}{34,98\%}         \\
        {500}                         & {500}     & {3} & {16}  & \textcolor{purple}{34,77\%}         \\
        {500}                         & {500}     & {6} & {2}   & \textcolor{purple}{34,38\%}         \\
        {1000}                        & {1000}    & {4} & {2}   & \textcolor{purple}{35,33\%}         \\
        {1000}                        & {1000}    & {6} & {100} & \textcolor{purple}{35,24\%}         \\
        \bottomrule
    \end{tabular}
    \caption{Selected results of the object identification games on the Dale-5 dataset with $e_s=100$ and $e_r=500$}
    \label{tab:results_discriminator_dale-5}
\end{table}

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{cccc|c}
        \toprule
                                      &           &     &       & \multicolumn{1}{c}{\textbf{CLEVR color}} \\  \cmidrule(lr){5-5}
        $h\_r$                        & $h\_s$    & $n$ & $|V|$ & \textbf{Accuracy}                        \\\midrule
        {100}                         & {1000}    & {3} & {10}  & {73,65\%}                                \\
        {100}                         & {1000}    & {3} & {16}  & {73,57\%}                                \\
        {500}                         & {500}     & {4} & {10}  & {73,57\%}                                \\
        {100}                         & {500}     & {3} & {16}  & {70,75\%}                                \\
        {100}                         & {1000}    & {4} & {50}  & {69,31\%}                                \\
        {1000}                        & {500}     & {1} & {16}  & {62,85\%}                                \\
        {100}                         & {500}     & {1} & {10}  & {57,12\%}                                \\
        {1000}                        & {500}     & {6} & {16}  & {54,69\%}                                \\
        {100}                         & {100}     & {4} & {2}   & {52,95\%}                                \\
        {1000}                        & {500}     & {1} & {100} & {52,82\%}                                \\
        {100}                         & {1000}    & {1} & {100} & {52,34\%}                                \\
        {100}                         & {500}     & {3} & {2}   & {51,48\%}                                \\
        {500}                         & {1000}    & {3} & {16}  & \textcolor{purple}{24,61\%}              \\
        {100}                         & {1000}    & {3} & {50}  & \textcolor{purple}{24,48\%}              \\
        {1000}                        & {1000}    & {4} & {100} & \textcolor{purple}{24\%}                 \\
        {500}                         & {1000}    & {1} & {2}   & \textcolor{purple}{23,96\%}              \\
        {1000}                        & {100}     & {6} & {10}  & \textcolor{purple}{23,87\%}              \\
        {500}                         & {1000}    & {6} & {50}  & \textcolor{purple}{23,87\%}              \\\midrule
        \multicolumn{4}{c|}{baseline} & {23,74\%}                                                          \\\midrule
        {500}                         & {500}     & {2} & {2}   & \textcolor{purple}{22,35\%}              \\
        \bottomrule
    \end{tabular}
    \caption{Selected results of the object identification games on the CLEVR color dataset with $e_s=100$ and $e_r=500$}
    \label{tab:results_discriminator_color}
\end{table}

>>>>

The language game is run on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
A random guess corresponds to 50\% in the \emph{Dale-2} dataset, 20\% in the \emph{Dale-5} dataset and 13\% in the \emph{CLEVR color} dataset.

@@

\begin{figure}[ht!]
    \centering
    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-2_vocab-size.png}
        \label{fig:learning-curve_discriminator_dale-2_vocab-size}
    }
    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-2_max-len.png}
        \label{fig:learning-curve_discriminator_dale-2_max-len}
    }
    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-5_vocab-size.png}
        \label{fig:learning-curve_discriminator_dale-5_vocab-size}
    }
    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-5_max-len.png}
        \label{fig:learning-curve_discriminator_dale-5_max-len}
    }
    \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_colour_vocab-size.png}
        \label{fig:learning-curve_discriminator_colour_vocab-size}
    }
    \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_colour_max-len.png}
        \label{fig:learning-curve_discriminator_colour_max-len}
    }
    \caption{Learning curves of all object identification games on each dataset. In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$. The baseline is marked in black.}
    \label{fig:learning-curves_discriminator}
\end{figure}

When looking at the results of the experiments, it can be seen that the agents are successfully discriminating the objects across all datasets.
Figure \ref{fig:learning-curves_discriminator} gives an overview of the learning process of the agents.
Each graph shows the learning curve of a configuration.
In the left column, every color stands for configurations with a certain vocabulary size while in the right column a color corresponds to a message length.
The rows show the three datasets.
First of all, it can be seen that there are many configurations that outperform the baseline across all datasets and the agents are able to solve the task together.
However, there are differences in the speed, the performance and the influence of vocabulary size and message length.

The agents can solve the task on \emph{Dale-2} dataset very fast.
Most of the configurations reach accuracies close to the final accuracy already after 2000 to 7000 games.
If the agents learn to discriminate the object, they do so during these first games.
After they reached these high accuracies, the learning is relatively unstable and can drop several times for up to 20\% points.
However, the agents can recover after some thousand games to the original performance.
The more games are played, the more stable the learning gets.
The influence of the vocabulary size and message length is not as clear, but smaller vocabulary sizes tend to help the agents to learn faster.
On the \emph{Dale-5} dataset, there is already a wider spread in performances and the learning is slower than on the \emph{Dale-2} dataset.
The agents learn fast up to an accuracy of around 50\%, but for most configurations it then slows down, and the accuracy improves until the cutoff after 128.000 games.
When the agents are given a vocabulary size of $|V|=2$, for several runs, the learning even stops after they reach around 50\% accuracy or progresses only very slowly.
With a vocabulary size of $|V|=100$, the agents reach the 50\% some thousand games after the remaining vocabulary sizes, but the learning goes on afterwards.
While there is no visible influence of the message length on the learning speed, all configurations that stay at an accuracy between 45\% and 65\% have a message length of $n=1$.
Interestingly, there is also one configuration that diverges from the baseline very late after almost 80.000 games.
Finally, on the \emph{CLEVR color} dataset, difference in learning speed are again more apparent.
Agents tend to learn the fastest with a vocabulary size of $|V|=2$, the accuracy however almost stagnates after the initial boost to around 40\%.
When agents are given a vocabulary of size $|V| \in \{10,16\}$, the initial learning takes place several thousand games later, but therefore reaches higher final accuracies.
As seen already for the \emph{Dale-5} dataset, most of the configurations converge in the first 20.000 games, but there are several converge much later.
An influence of the message length on the speed is again not visible.
% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge. As David pointed out it may be that convergence is unstable and changes happen in between.
% DK: done

Tables \ref{tab:results_discriminator_dale-2} to \ref{tab:results_discriminator_color} give an overview over the final accuracy scores after 128.000 games of selected configurations.
For each dataset there are many configurations that outperform the baseline models by a high margin.
However, there are still clear differences between the accuracy scores of each dataset.

\begin{table}[ht]
    \centering
    \begin{tabular}{cccc|c}
        \toprule
                                      &           &     &       & \multicolumn{1}{c}{\textbf{Dale-2}} \\  \cmidrule(lr){5-5}
        $h_r$                         & $h_s$     & $n$ & $|V|$ & \textbf{Accuracy}                   \\\midrule
        {100}                         & {100}     & {2} & {10}  & {99,96\%}                           \\
        {100}                         & {1000}    & {2} & {10}  & {99,96\%}                           \\
        {1000}                        & {100}     & {3} & {10}  & {99,87\%}                           \\
        {500}                         & {100}     & {3} & {100} & {99,83\%}                           \\
        {100}                         & {100}     & {6} & {50}  & {99,83\%}                           \\
        {100}                         & {500}     & {3} & {16}  & {99,7\%}                            \\
        {1000}                        & {100}     & {6} & {16}  & {99,61\%}                           \\
        {100}                         & {100}     & {6} & {100} & {99,57\%}                           \\
        {100}                         & {500}     & {1} & {100} & {99,52\%}                           \\
        {100}                         & {100}     & {1} & {50}  & {99,35\%}                           \\
        {500}                         & {100}     & {1} & {2}   & {99,26\%}                           \\
        {100}                         & {500}     & {1} & {50}  & {99,26\%}                           \\
        {500}                         & {500}     & {1} & {100} & {99,26\%}                           \\
        {100}                         & {500}     & {3} & {100} & {99,83\%}                           \\
        {500}                         & {1000}    & {1} & {16}  & {98\%}                              \\
        {100}                         & {500}     & {1} & {10}  & {96,61\%}                           \\
        {100}                         & {1000}    & {1} & {100} & {99,44\%}                           \\
        {500}                         & {500}     & {6} & {2}   & {95,92\%}                           \\
        {100}                         & {1000}    & {4} & {2}   & {94,88\%}                           \\
        {100}                         & {1000}    & {6} & {10}  & \textcolor{purple}{63,15\%}         \\
        {1000}                        & {500}     & {2} & {2}   & \textcolor{purple}{62,98\%}         \\\midrule
        \multicolumn{4}{c|}{baseline} & {62,72\%}                                                     \\\midrule
        {100}                         & {1000}    & {3} & {2}   & \textcolor{purple}{61,2\%}          \\
        \bottomrule
    \end{tabular}
    \caption{Selected results of the object identification games on the Dale-2 dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
    \label{tab:results_discriminator_dale-2}
\end{table}

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{cccc|c}
        \toprule
                                      &           &     &       & \multicolumn{1}{c}{\textbf{Dale-5}} \\  \cmidrule(lr){5-5}
        $h_r$                         & $h_s$     & $n$ & $|V|$ & \textbf{Accuracy}                   \\\midrule
        {1000}                        & {500}     & {3} & {50}  & {91,75\%}                           \\
        {100}                         & {100}     & {6} & {10}  & {91,02\%}                           \\
        {100}                         & {100}     & {3} & {10}  & {90,76\%}                           \\
        {100}                         & {100}     & {4} & {16}  & {89,76\%}                           \\
        {100}                         & {100}     & {2} & {10}  & {88,85\%}                           \\
        {500}                         & {1000}    & {1} & {100} & {87,11\%}                           \\
        {1000}                        & {1000}    & {6} & {50}  & {85,07\%}                           \\
        {500}                         & {1000}    & {1} & {10}  & {84,46\%}                           \\
        {500}                         & {500}     & {1} & {50}  & {83,29\%}                           \\
        {100}                         & {100}     & {1} & {10}  & {80,86\%}                           \\
        {500}                         & {100}     & {1} & {100} & {76,74\%}                           \\
        {500}                         & {500}     & {4} & {2}   & {77,78\%}                           \\
        {1000}                        & {1000}    & {3} & {2}   & {75,78\%}                           \\
        {100}                         & {100}     & {3} & {2}   & {74\%}                              \\
        {500}                         & {500}     & {1} & {2}   & {63,41\%}                           \\\midrule
        \multicolumn{4}{c|}{baseline} & {35,07\%}                                                     \\\midrule
        {500}                         & {1000}    & {2} & {100} & \textcolor{purple}{34,98\%}         \\
        {500}                         & {500}     & {3} & {16}  & \textcolor{purple}{34,77\%}         \\
        {500}                         & {500}     & {6} & {2}   & \textcolor{purple}{34,38\%}         \\
        {1000}                        & {1000}    & {4} & {2}   & \textcolor{purple}{35,33\%}         \\
        {1000}                        & {1000}    & {6} & {100} & \textcolor{purple}{35,24\%}         \\
        \bottomrule
    \end{tabular}
    \caption{Selected results of the object identification games on the Dale-5 dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
    \label{tab:results_discriminator_dale-5}
\end{table}

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{cccc|c}
        \toprule
                                      &           &     &       & \multicolumn{1}{c}{\textbf{CLEVR color}} \\  \cmidrule(lr){5-5}
        $h\_r$                        & $h\_s$    & $n$ & $|V|$ & \textbf{Accuracy}                        \\\midrule
        {100}                         & {1000}    & {3} & {10}  & {73,65\%}                                \\
        {100}                         & {1000}    & {3} & {16}  & {73,57\%}                                \\
        {500}                         & {500}     & {4} & {10}  & {73,57\%}                                \\
        {100}                         & {500}     & {3} & {16}  & {70,75\%}                                \\
        {100}                         & {1000}    & {4} & {50}  & {69,31\%}                                \\
        {1000}                        & {500}     & {1} & {16}  & {62,85\%}                                \\
        {100}                         & {500}     & {1} & {10}  & {57,12\%}                                \\
        {1000}                        & {500}     & {6} & {16}  & {54,69\%}                                \\
        {100}                         & {100}     & {4} & {2}   & {52,95\%}                                \\
        {1000}                        & {500}     & {1} & {100} & {52,82\%}                                \\
        {100}                         & {1000}    & {1} & {100} & {52,34\%}                                \\
        {100}                         & {500}     & {3} & {2}   & {51,48\%}                                \\
        {500}                         & {1000}    & {3} & {16}  & \textcolor{purple}{24,61\%}              \\
        {100}                         & {1000}    & {3} & {50}  & \textcolor{purple}{24,48\%}              \\
        {1000}                        & {1000}    & {4} & {100} & \textcolor{purple}{24\%}                 \\
        {500}                         & {1000}    & {1} & {2}   & \textcolor{purple}{23,96\%}              \\
        {1000}                        & {100}     & {6} & {10}  & \textcolor{purple}{23,87\%}              \\
        {500}                         & {1000}    & {6} & {50}  & \textcolor{purple}{23,87\%}              \\\midrule
        \multicolumn{4}{c|}{baseline} & {23,74\%}                                                          \\\midrule
        {500}                         & {500}     & {2} & {2}   & \textcolor{purple}{22,35\%}              \\
        \bottomrule
    \end{tabular}
    \caption{Selected results of the object identification games on the CLEVR color dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
    \label{tab:results_discriminator_color}
\end{table}
